---
title: "Data Visualisation Notebook"
output: html_notebook
---

# R for Data Science 

## Chapter 3: Data Visualisation

### 3.1 Introduction

#### 3.1.1 Prerequisites

```{r}
library(tidyverse)
```

### 3.2 First Steps

#### 3.2.1 The mpg dataframe

```{r}
mpg
```

#### 3.2.2 Creating a ggplot

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

#### 3.2.3 A graphing template

ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
  
#### 3.2.4 Exercises
  
```{r}
# Nothing appears
ggplot(data = mpg)
```

```{r}
# How many rows are in mpg? 
nrow(mpg)

# How many columns?
ncol(mpg)
```

What does the drv variable describe? Read the help for ?mpg to find out.
```{r}
?mpg
# drv: f = front-wheel drive, r = rear wheel drive, 4 = 4wd
```

```{r}
# a scatterplot of hwy vs cyl.
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl))
```
```{r}
# What happens if you make a scatterplot of class vs drv? Why is the plot not useful?
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = class, y = drv))
# This plot is not useful because the values of both axes are non-numerical
```

### 3.3 Aesthetic mappings

  aesthetic - a visual property of the objects in your plot
```{r}
# plots displ vs hwy with class differentiated with color (legend included)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

```{r}
# plots displ vs hwy with class differentiated by size (legend included)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))

# problematic because size is not easily differentiated visually
```

```{r}
# Left
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))

# Right
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))

```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

#### 3.3.1 Exercises

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
# color = "blue" fulfills the class argument in aes() function, instead of a argument within geom_point() function (which is where the color of the points will change). 
```

```{r}
# Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?
?mpg
mpg # when you run mpg if the values are a number they are continuous, if the values are names, they are categorical
```
categorical: manufacturer model, trans, drv, fl, class
continuous: displ, year, cyl, cty, hwy

```{r}
# Map a continuous variable to color, size, and shape
# color
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = cyl))
# size: 
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = year))
# shape: error cannot map continuous variable to shape
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = cty))
```
Continuous vs categorical plotting using shape, size, and color:
  continuous plot of color has a gradient bar instead of categorical plot of color has diff colored dots
  continuous plot of size has the same dots as categorical plot
  continuous plot of shape cannot be plotted 

What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)
```{r}
?geom_point
# stroke aesthetic modifies size of the points
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = year), stroke = 5)
```
What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy, color = displ < 5))
# instead of differentiating the values of all displ options, the colors differ between what is less than 5 (True) and what is greater than or equal to 5 (False)
```

### 3.4 Common Problems

```{r}
ggplot(data = mpg) 
+ geom_point(mapping = aes(x = displ, y = hwy))
# does not work
# should be 
# ggplot(data = mpg) + 
#   geom_point(mapping = aes(x = displ, y = hwy))
```
### 3.5 Facets 
facet_wrap() - subplots that each display one subplot of the data
```{r}
# plots data from mpg
# specifically displ as x axis and hwy as y axis
# separate points into subplot based on class, in 2 rows
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

```{r}
# plots data from mpg
# specifically displ as x axis and hwy as y axis
# separate points into subplot based on drv (top x axis), and cyl (right y axis)
# note: facet_grid() vs facet_wrap()
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```

```{r}
# plots data from mpg
# specifically displ as x axis and hwy as y axis
# separate points into subplot based on cyl (top x axis), 2 rows
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ cyl)

# plots data from mpg
# specifically displ as x axis and hwy as y axis
# separate points into subplot based on cyl (top x axis), 1 row
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(. ~ cyl)

# plots data from mpg
# specifically displ as x axis and hwy as y axis
# separate points into subplot based on cyl (right y axis)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(cyl ~ .)
```

#### 3.5.1 Excersies



    What happens if you facet on a continuous variable?
```{r}
# facet on a continuous variable results in a lot of subplots (not necessarily helpful)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ cty)
```

    What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?
```{r}
# cyl means number of cylinders
# drv can be f = front-wheel drive, r = rear wheel drive, 4 = 4wd
?mpg

# plots data from mpg
# specifically displ as x axis and hwy as y axis
# separate points into subplot based on drv (top x axis), and cyl (right y axis)
# note empty subplots: 4 cyl-rear wheel drive, 5 cyl-4 wheel drive, 5 cyl-rear wheel drive
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)

# plots data from mpg
# specifically drv as x axis and cyl as y axis
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = drv, y = cyl))
```
  The intersections are blank on both graphs, which means that as a combination, 4 cyl-rear wheel drive, 5 cyl-4 wheel drive, 5 cyl-rear wheel drive do not exist. Also there is no car that has 7 cylinders

    What plots does the following code make? What does . do?
  . is a placeholder so that the specified variable can remain in either 1 row or 1 column
```{r}
# plots data from mpg
# specifically displ as x axis and hwy as y axis
# separate points into subplot based on drv (top x axis)
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) +
      facet_grid(drv ~ .)

# plots data from mpg
# specifically displ as x axis and hwy as y axis
# separate points into subplot based on cyl (right y axis)
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) +
      facet_grid(. ~ cyl)
```
    
    Take the first faceted plot in this section:
```{r}
# plots data from mpg
# specifically displ as x axis and hwy as y axis
# separate points into subplot based on drv (top x axis)
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) + 
      facet_wrap(~ class, nrow = 2)
```
    What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?
- The advantages is that it would properly plot overlapping points of different class (specified variable). The disadvantage is that it is hard to see the overall relationship between all of the points of different class. The larger the dataset would make more sense to use facet if the points were overlapping and not spread out in their own clusters in one plot

    Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?
```{r}
?facet_wrap
```
- nrow sets the number of rows that the subplots will appear in
- ncol sets the number of columns the subplots will appear in
- options that control the layout of individual panels: scales, shrink, as.table, switch, drop, dir, strip.position
- facet_grid() has no nrow or ncol arguments because each axis is taken up by 2 different variables. To achieve a grid like structure, you need each axis to occupy 1 row or 1 column (each axis must have 1 dim = 1)

    When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?
- Because there is a fixed width for the columns in which the rows are not limited (like a sheet of paper, there is more room going down the paper than left to right)

### 3.6 Geometric Objects
```{r}
# same variables on the axes but different geoms (geom is the geometrical object used to represent data)
# left: geom_point() plots points
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# right: geom_smooth() plots a smooth line and a gray area for error(?)
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
# can add a 3rd variable like geom_point() but now the argument is specifically linetype
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

```{r}
# guessed at code to recreate the plot 
# note: order matters
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point(mapping = aes(x = displ, y = hwy, color = drv))

# some points are darker or covered up by lines now 
# this is the actual graph 
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_smooth(mapping = aes(x = displ, y = hwy, color = drv))
```

```{r}
# Left: the original plot using geom_smooth without a third variable
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
# Middle: uses group // does not distinguish between the different groups            
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
# Right: distinguishes drv by color but does not show legend
# normally would automatically show legend unless specified show.legend= FALSE
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )
```

```{r}
# plots points then smooth line (like a couple chunks ago)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
# simplifies the code so that you are only changing one variable in ggplot() even though it is the same as the previous chunk
# this is important because you can have errors through referencing multiple variables in different places
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()
```

```{r}
# adds in color to just geom_point()
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()

# gets a lot of errors
# tried to map color for both point and smooth ~ not super helpful for visualizing data
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) + 
  geom_point() + 
  geom_smooth()
```

```{r}
# still using color for class in geom_point() only
# filtering only subcompact class of the smoothed data

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)

# se argument is the gray area around the line (confidence interval), TRUE by default
?geom_smooth
```

#### 3.6.1 Exercises

    What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
```{r}
# line chart
ggplot(data = mpg) + 
  geom_line(mapping = aes(x = displ, y = hwy, color = class))

# boxplot
ggplot(data = mpg) + 
  geom_boxplot(mapping = aes(x = displ, y = hwy, color = class))

# histogram
ggplot(data = mpg) + 
  geom_histogram(mapping = aes(x = displ))

# area chart
ggplot(data = mpg) + 
  geom_area(mapping = aes(x = displ, y = hwy, color = class))

```

    Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
- I predict that this code will plot displ and hwy as points, differing the drv by color. The smooth lines will also be divided drv by color and will not show the gray area around the lines

```{r}
# Prediction = TRUE
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

    What does show.legend = FALSE do? What happens if you remove it?
    Why do you think I used it earlier in the chapter?
- show.legend = FALSE does not display the legend. 
- If you remove this, it defaults to show.legend = TRUE and shows the legend
- It was used earlier in the chapter because the other graphs did not have a legend and having a legend slightly shrinks the x axis (visual evenness)

    What does the se argument to geom_smooth() do?
- se argument is in reference to the gray area near the line. It signifies the confidence interval and the se argument specifies if it is visible on the graph (TRUE/FALSE)
    
    Will these two graphs look different? Why/why not?
- these graphs will not look different because if the mapping is specified in ggplot() then it applies to all the functions following (+) 
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

    Recreate the R code necessary to generate the following graphs.
    
```{r}
# top left: no confidence interval, both a line and points
# x axis is displ and y axis is hwy
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

# top right: same as the previous graph, except the line is split into 3 (we've seen this previously as drv)
# drv line is split as group because there is no discerning legend or color (linetype etc)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(mapping = aes(group = drv), se = FALSE)

# mid left: both the lines and points have discerning color by drv, no confidence interval visible still
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth( se = FALSE)

# mid right: color difference for points, just 1 line though. 
# specify color in geom_point() but not in geom_smooth()
# still no confidence interval (se = FALSE)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv)) + 
  geom_smooth(se = FALSE)

# bottom left: points different through color (drv), line different through line type (drv)
# no confidence interval
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv)) + 
  geom_smooth(mapping = aes(linetype = drv), se = FALSE)

# bottom right: order matters
# white points behind the color points separated by drv
# white points are visible from behind color points, must be larger - stroke = 2
# specify color and stroke for first geom_point()
# mapping for drv, separate by color for second geom_point()
# if done in the wrong order, white dots, no color
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(color = "white", stroke = 2) +
  geom_point(mapping = aes(color = drv)) 
```
### 3.7 Statistical Transformations
```{r}
# plots cut as x axis and freq/count as y axis
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
```

```{r}
# same plot different command; stat_count()
ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))
```

```{r}
# again, same concept, different method, different data too
demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)

ggplot(data = demo) +
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
```

```{r}
# using proportion instead of default mapping
# note the y axis (and compare to the previous diamond cut plots )
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
```

```{r}
# maps specific x (cut) and y (depth)
# specifies the y axis bounds using data information
# plots the median of each x axis value 
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

```{r}
# lots of stats to use in graphing - check em out
?stat_bin
```

#### 3.7.1 Exercises

    What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

```{r}
?stat_summary
# default connection between geom_histogram/geom_freqpoly
?geom_crossbar

# this is as close as I got to the original graph using geom_line()
ggplot(data = diamonds, mapping = aes(x = cut, y = depth)) + 
  geom_line(ymin = min(diamonds$depth), ymax = max(diamonds$depth))
```


    What does geom_col() do? How is it different to geom_bar()?
```{r}
# plots cut as x axis and freq/count as y axis
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))

# Need to specify a y axis - err
ggplot(data = diamonds) + 
  geom_col(mapping = aes(x = cut))
```

    Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?

    What variables does stat_smooth() compute? What parameters control its behaviour?

    In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?

```{r}
    ggplot(data = diamonds) + 
      geom_bar(mapping = aes(x = cut, y = ..prop..))
    ggplot(data = diamonds) + 
      geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))
```

```{r}
    ggplot(data = diamonds) + 
      geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
    ggplot(data = diamonds) + 
      geom_bar(mapping = aes(x = cut, fill = color, y = ..prop.., group = 1))
```


### 3.8 Position adjustments

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
```

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(alpha = 1/5, position = "identity")
ggplot(data = diamonds, mapping = aes(x = cut, colour = clarity)) + 
  geom_bar(fill = NA, position = "identity")
```

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
```

#### 3.8.1 Exercises

    What is the problem with this plot? How could you improve it?
```{r}
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
      geom_point()
```

    What parameters to geom_jitter() control the amount of jittering?

    Compare and contrast geom_jitter() with geom_count().

    What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.

### 3.9 Coordinate systems

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
```

```{r}
nz <- map_data("nz")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_quickmap()
```

```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()
```

#### 3.9.1 Exercises

    Turn a stacked bar chart into a pie chart using coord_polar().

    What does labs() do? Read the documentation.

    What’s the difference between coord_quickmap() and coord_map()?

    What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```

### 3.10 The layered grammar of graphics

ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>
  
  