---
title: "NLP Notebook"
output: html_notebook
---

# Natural Language Processing

### Installing the Necessary Libraries

```{r}
library(NLP)
library(openNLP)
library(RWeka)
library(rJava)
library(magrittr)
```


I was warned it would get angry

## Basic Tokenization

```{r}
# made the folder path within the homework folder
bio <- readLines("data/nlp/anb-jarena-lee.txt")
print(bio)
```

```{r}
# This removes all new lines from the character vector (one line now)
bio <- paste(bio, collapse = " ")
print(bio)
```
## Sentence and Word Annotations

```{r}
# changes the character vector to a string object
bio <- as.String(bio)
```
```{r}
# Annotators created from Annotation functions
# Annotators saved to variables
# word annotator
word_ann <- Maxent_Word_Token_Annotator()
# sentence annotator
sent_ann <- Maxent_Sent_Token_Annotator()
```

```{r}
# apply word and sentence annotators to bio string object
bio_annotations <- annotate(bio, list(sent_ann, word_ann))
```

```{r}
# information about bio_annotations
class(bio_annotations)
```

```{r}
# The first 6 lines of bio_annotations
head(bio_annotations)
```


Associated metadata using meta= argument

```{r}
# AnnotatedPlainTextDocument - combining the document and the annotations
bio_doc <- AnnotatedPlainTextDocument(bio, bio_annotations)
```
View(bio_doc) does not reveal much

```{r}
# to get the sentences (but just the first two) separated into words and punctuation
sents(bio_doc) %>% head(2)
```

```{r}
# the first 10 words in bio_doc
words(bio_doc) %>% head(10)
```

## Annotation people and places

technique: named entity recognition (NER) 

```{r}
# These below are all functions to get people, locations, and organization ~ curious how the function is set up for "organization"
person_ann <- Maxent_Entity_Annotator(kind = "person")
location_ann <- Maxent_Entity_Annotator(kind = "location")
organization_ann <- Maxent_Entity_Annotator(kind = "organization")
```

```{r}
# creating an annotated plain text document 
# this line combines all of the annotators into one list / variable
pipeline <- list(sent_ann,
                 word_ann,
                 person_ann,
                 location_ann,
                 organization_ann)
# bio_annotations <- applies all annotators to string object
bio_annotations <- annotate(bio, pipeline)
# bio_doc creating an annotated plain text document 
bio_doc <- AnnotatedPlainTextDocument(bio, bio_annotations)
```

```{r}
# easy function for extracting entities from an Annotated Plain Text Document (we will use this in the next code chunk)
entities <- function(doc, kind) {
  s <- doc$content
  a <- annotation(doc) 
  #originally a <- annotations(doc)[[1]]
  if(hasArg(kind)) {
    k <- sapply(a$features, `[[`, "kind")
    s[a[k == kind]]
  } else {
    s[a[a$type == "entity"]]
  }
}
```

```{r}
# use entities function from the previous code chunk
# separating out "person" entity
entities(bio_doc, kind = "person")
```
^ originally this line did not work (output: character(0)) and there were a couple of errors within the original function 
- annotations(doc) is not a function
  I found this out through ?annotation in the console
- annotation(doc)[[1]] also finds nothing
  in comparison with the results that you find in the tutorial, I checked the output of annotation(bio_doc) which gives a long list and at the end, shows the kind= people, location, and organization (which is surely what we are looking for)
```{r}
# shows the annotations of the annotated plain text document
annotation(bio_doc)
```
  
And if you stick [[1]]  at the end of that command, it only takes the first line of the annotations list (which does not have kind = person, location, organization)
```{r}
annotation(bio_doc)[[1]]
```  

```{r}
# extracted locations from the bio_doc variable
entities(bio_doc, kind = "location")
```

```{r}
# extracted organizations from the bio_doc variable
entities(bio_doc, kind = "organization")
```

# Named Entity Recognition in a small corpus

```{r}
# gets all of the filenames using the glob naming conventions *.txt
filenames <- Sys.glob("data/itinerants/*.txt")
filenames
```

```{r}
# names assigning to the list (clean up)
texts <- filenames %>%
  lapply(readLines) %>%
  lapply(paste0, collapse = " ") %>%
  lapply(as.String)

names(texts) <- basename(filenames)

str(texts, max.level = 1)
```

```{r}
# a function that annotates text and then creates a AnnotatedPlainTextDocument
annotate_entities <- function(doc, annotation_pipeline) {
  print("Err")
  annotations <- annotate(doc, annotation_pipeline)
  print("eer2")
  AnnotatedPlainTextDocument(doc, annotations)
}
```

```{r}
# creates a list of the different annotators: word, sentence, person, location 
itinerants_pipeline <- list(
  Maxent_Sent_Token_Annotator(),
  Maxent_Word_Token_Annotator(),
  Maxent_Entity_Annotator(kind = "person"),
  Maxent_Entity_Annotator(kind = "location")
)
```
Below - do not run
```{r}
texts_annotated <- texts %>%
  lapply(annotate_entities, itinerants_pipeline)
```

```{r}
places <- texts_annotated %>%
  lapply(entities, kind = "location")

people <- texts_annotated %>%
  lapply(entities, kind = "person")
```

```{r}
places %>%
  sapply(length)
```

```{r}
places %>%
  lapply(unique) %>%
  sapply(length)
```

```{r}
people %>%
  sapply(length)
```

```{r}
people %>%
  lapply(unique) %>%
  sapply(length)
```

```{r}
library(ggmap)
```

```{r}
all_places <- union(places[["pratt-parley.txt"]], places[["cartwright-peter.txt"]]) %>% union(places[["lee-jarena.txt"]])
```

