---
title: "R Notebook"
output: html_notebook
---

### Trying it on my own texts


Initially I thought to do some exploring through code since it seemed more efficient than searching through the meta-data myself

```{r}
library(gutenbergr)
library(dplyr)
gutenberg_metadata
```

I still haven't gotten use to calling the functions in R so I looked through the variable  and clicked on it within the top right corner of R Studio and looked through that way. I'll list the steps down so that I can come back and see if I can replicate this in code.

1. Search for Lewis Carroll texts, find author ID (7)
2. Search for Frank L. Baum texts, find author ID (42)
3. Search for A. A. Milne texts, find author ID (730)
```{r}
# I was hoping this line just downloaded the texts corresponding to the specific author IDs
gutenberg_download(730, mirror = "http://gutenberg.pglaf.org/")
#i was wrong
```
This chunk (below) lists the works under each author and stores it into the respective variale 
```{r}
lewis <- gutenberg_metadata %>% filter(author == "Caroll, Lewis")
baum <- gutenberg_metadata %>% filter(author == "Baum, L. Frank (Lyman Frank)")
milne <- gutenberg_metadata %>% filter(author == "Milne, A. A. (Alan Alexander)")
```

```{r}
carroll <- gutenberg_works(author == "Carroll, Lewis") %>%
  gutenberg_download(meta_fields = "title", mirror = "http://gutenberg.pglaf.org/")
baum <- gutenberg_works(author == "Baum, L. Frank (Lyman Frank)") %>%
  gutenberg_download(meta_fields = "title", mirror = "http://gutenberg.pglaf.org/")
milne <- gutenberg_works(author == "Milne, A. A. (Alan Alexander)") %>%
  gutenberg_download(meta_fields = "title", mirror = "http://gutenberg.pglaf.org/")
```

