---
title: "Week 3: Experiments - Elaborations"
output: html_notebook
---

## Objectives and Output

For more on how to manipulate tidy texts and dataframes: R for Data Science, chapter 5: http://r4ds.had.co.nz/ (Links to an external site.)

For an introduction to R programming: The Art of R Programming, chapters 1 and 2: https://ebookcentral.proquest.com/lib/fsu/detail.action?docID=1137514


## R for Data Science

### Chapter 5: Data Transformation

#### 5.1 Introduction

how to transform your data using the dplyr package (and a new dataset on flights)

##### 5.1.1 Prerequisites
At this point, I was having trouble installing the tidyverse package. so I had to update my R version 

To do so, I went http://mercury.webster.edu/aleshunas/R_learning_infrastructure/Updating%20R%20and%20RStudio.html for update instructions. 

*R version 3.2.3 (2015-12-10) -- "Wooden Christmas-Tree"*
^ This is my current version. As you can see, it was released around 2015. The current version available is R version 3.5.1 (Feather Spray) released 2018-07-02.

So even thoughI tried to install a new R package, it seems that the packages dpylr and tidytext provide the same function.
```{r}
library(nycflights13)
#library(tidyverse)
library(tidytext)
library(dplyr)
```

(Conflicting function calls)
If you want to use the base version of these functions after loading dplyr, you’ll need to use their full names: stats::filter() and stats::lag().

##### 5.1.2 nycflights13

```{r}
flights
```

##### 5.1.3 dplyr basics

filter()
arrange()
select()
mutate()
summarize()

group_by()

#### 5.2 filter rows with filter()

```{r}
filter(flights, month == 1, day == 1)
```


```{r}
jan1 <- filter(flights, month == 1, day == 1)
```


```{r}
(dec25 <- filter(flights, month == 12, day == 25))
```

##### 5.2.1 Comparisons
```{r}
sqrt(2) ^ 2 == 2
1 / 49 * 49 == 1
```


```{r}
near(sqrt(2) ^ 2,  2)
near(1 / 49 * 49, 1)
```

##### 5.2.2 Logical Operators
```{r}
filter(flights, month == 11 | month == 12)
```

```{r}
nov_dec <- filter(flights, month %in% c(11, 12))
```

```{r}
filter(flights, !(arr_delay > 120 | dep_delay > 120))
filter(flights, arr_delay <= 120, dep_delay <= 120)
```

##### 5.2.3 Missing Values

```{r}
NA > 5
10 == NA
NA + 10
NA / 2
```

```{r}
NA == NA
```

```{r}
# Let x be Mary's age. We don't know how old she is.
x <- NA

# Let y be John's age. We don't know how old he is.
y <- NA

# Are John and Mary the same age?
x == y
```

```{r}
is.na(x)
```

```{r}
df <- tibble(x = c(1, NA, 3))
filter(df, x > 1)

filter(df, is.na(x) | x > 1)
```

##### 5.2.4 Exercises
Find all flights
```{r}
# Had an arrival delay of two or more hours
arrDelay2 <- filter(flights, arr_delay >= 2)
```

```{r}
# Flew to Houston (IAH or HOU)
filter(flights, dest == "IAH"| dest == "HOU")
```

```{r}
# Were operated by United, American, or Delta
filter(flights, carrier %in% c("UA", "AA", "DL"))
```

```{r}
# Departed in summer (July, August, and September)
filter(flights, month %in% c(7, 8, 9))
```

```{r}
# Arrived more than two hours late, but didn’t leave late

# Were delayed by at least an hour, but made up over 30 minutes in flight
# Departed between midnight and 6am (inclusive)

```

Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)


## Text Mining with R

### Ch 2: Sentient Analysis with tidy Data

### Ch 3: Analyzing word and document frequency: tf-idf

Examples from this chapter were already done in the previous homework assignment
