---
title: "Week 3: Experiments - Elaborations"
output: html_notebook
---

## Objectives and Output

For more on how to manipulate tidy texts and dataframes: R for Data Science, chapter 5: http://r4ds.had.co.nz/ (Links to an external site.)

For an introduction to R programming: The Art of R Programming, chapters 1 and 2: https://ebookcentral.proquest.com/lib/fsu/detail.action?docID=1137514


## R for Data Science

### Chapter 5: Data Transformation

#### 5.1 Introduction

how to transform your data using the dplyr package (and a new dataset on flights)

##### 5.1.1 Prerequisites
At this point, I was having trouble installing the tidyverse package. so I had to update my R version 

To do so, I went http://mercury.webster.edu/aleshunas/R_learning_infrastructure/Updating%20R%20and%20RStudio.html for update instructions. 

*R version 3.2.3 (2015-12-10) -- "Wooden Christmas-Tree"*
^ This is my current version. As you can see, it was released around 2015. The current version available is R version 3.5.1 (Feather Spray) released 2018-07-02.

So even thoughI tried to install a new R package, it seems that the packages dpylr and tidytext provide the same function.
```{r}
library(nycflights13)
library(tidyverse)
library(tidytext)
library(dplyr)
```

(Conflicting function calls)
If you want to use the base version of these functions after loading dplyr, you’ll need to use their full names: stats::filter() and stats::lag().

##### 5.1.2 nycflights13

```{r}
flights
```

##### 5.1.3 dplyr basics

filter()
arrange()
select()
mutate()
summarize()

group_by()

#### 5.2 filter rows with filter()

```{r}
filter(flights, month == 1, day == 1)
```


```{r}
jan1 <- filter(flights, month == 1, day == 1)
```


```{r}
(dec25 <- filter(flights, month == 12, day == 25))
```

##### 5.2.1 Comparisons
```{r}
sqrt(2) ^ 2 == 2
1 / 49 * 49 == 1
```


```{r}
near(sqrt(2) ^ 2,  2)
near(1 / 49 * 49, 1)
```

##### 5.2.2 Logical Operators
```{r}
filter(flights, month == 11 | month == 12)
```

```{r}
nov_dec <- filter(flights, month %in% c(11, 12))
```

```{r}
filter(flights, !(arr_delay > 120 | dep_delay > 120))
filter(flights, arr_delay <= 120, dep_delay <= 120)
```

##### 5.2.3 Missing Values

```{r}
NA > 5
10 == NA
NA + 10
NA / 2
```

```{r}
NA == NA
```

```{r}
# Let x be Mary's age. We don't know how old she is.
x <- NA

# Let y be John's age. We don't know how old he is.
y <- NA

# Are John and Mary the same age?
x == y
```

```{r}
is.na(x)
```

```{r}
df <- tibble(x = c(1, NA, 3))
filter(df, x > 1)

filter(df, is.na(x) | x > 1)
```

##### 5.2.4 Exercises
Find all flights
```{r}
# Had an arrival delay of two or more hours
arrDelay2 <- filter(flights, arr_delay >= 120)
```

```{r}
# Flew to Houston (IAH or HOU)
filter(flights, dest == "IAH"| dest == "HOU")
```

```{r}
# Were operated by United, American, or Delta
filter(flights, carrier %in% c("UA", "AA", "DL"))
```

```{r}
# Departed in summer (July, August, and September)
filter(flights, month %in% c(7, 8, 9))
```

```{r}
# Arrived more than two hours late, but didn’t leave late
filter(flights, arr_delay > 120 & dep_delay <= 0)
```

```{r}
# Were delayed by at least an hour, but made up over 30 minutes in flight
filter(flights, dep_delay >= 60 & sched_arr_time - arr_time > 30)
```


```{r}
# Departed between midnight and 6am (inclusive)
filter(flights, dep_time >= 0  & dep_time <= 600)
```

Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

  between() uses a upper and lower bound to note which of the list lies between these values
  
```{r}
# Departed between midnight and 6am (inclusive) 
# using between function
filter(flights, between(dep_time, 0, 600))
```


How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
# number of flights with missing departure times 
# this is for flights that have been cancelled
nrow(filter(flights, is.na(dep_time)))
```

Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

```{r}
# results in 1
# anything to the power of 0 will always be 1
NA^0
```
```{r}
# results in TRUE
# anything with TRUE in an or statement will always be true
NA | TRUE
```
```{r}
# Since the two possibilities will either be true or false and if there is one false in an AND statement, it will always be false
FALSE & NA
```

#### 5.3 Arrange rows with arrange()

```{r}
# Ascending in year then month and then day
arrange(flights, year, month, day)
```

```{r}
# desc(x) is the x column in descending order 
arrange(flights, desc(dep_delay))
```

```{r}
# Purpose: missing values (NA) are always sorted at the end 
df <- tibble(x = c(5, 2, NA))
arrange(df, x)
# 2, 5, NA
arrange(df, desc(x))
# 5, 2, NA
```


##### 5.3.1 Exercises

```{r}
# how to use arrange() to sort all missing values to the start
arrange(flights, !is.na(dep_time), dep_time)

# i still don't really understand why this works. only that it does (somehow)
# oh got it now. second parameter states is not NA in Departure time (which will sort true or false -- and false is alphabetically before true) then sorts the numbers in ascending order (because the entries with numbers are all equally true [is not NA] )
```

```{r}
# Sort flights to find the most delayed flights. 
# Not sure whether this means departure or arrival so I included both
# desc because the larger the value, the bigger the delay
arrange(flights, desc(dep_delay), desc(arr_delay))
```

```{r}
# Find the flights that left earliest.
# early = negative departure delay time (sort ascending)
arrange(flights, dep_delay)
```

```{r}
# Sort flights to find the fastest flights
# by not overcomplicating things, fastest I assume to mean shortest air time (ascending order)
arrange(flights, air_time)
```

```{r}
# which flights travelled the longest
# since we've already covered time, this I assume to mean distance (which is a column) and the longer the distance, the longer the travel (descending order)
arrange(flights, desc(distance))
```

```{r}
# which flights travelled the shortest
# and the opposite from the previous chunk, ascending order for distance of the shortest flight
arrange(flights, distance)
```


#### 5.4 Select columns with select()

```{r}
# selecting specific columns
select(flights, year, month, day)
```

```{r}
# select using : which means all columns in between
select(flights, year:day)
```

```{r}
# select all columns except for those specified (-)
select(flights, -(year:day))
```

helper functions

starts_with("abc"): matches names that begin with “abc”.

ends_with("xyz"): matches names that end with “xyz”.

contains("ijk"): matches names that contain “ijk”.

matches("(.)\\1"): selects variables that match a regular expression. This one matches any variables that contain repeated characters. You’ll learn more about regular expressions in strings.

num_range("x", 1:3): matches x1, x2 and x3.

select() has rename capabilities but its finicky at best, just use rename() instead

```{r}
rename(flights, tail_num = tailnum)
```


##### 5.4.1 Exercises

```{r}
# select dep_time, dep_delay, arr_time, and arr_delay from flights.
select(flights, dep_time, dep_delay, arr_time, arr_delay)
select(flights, dep_time:arr_delay, -sched_dep_time, -sched_arr_time)

```

```{r}
# same variable multiple times
select(flights, year, year)
# shows only one column
```

```{r}
# one_of() is a select helper function for select()
# it uses a character vector to select the columns specified in the character vector

# these lines specify a character vector "vars" and select() will output a table of those variables listed in "vars"
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars))
```

```{r}
# what you think it does: if it contains TIME (case sensitive or at the beginning of the column name) it will keep that column

# what it actually does: if it contains time anywhere (case insensitive) in the column name it will keep that column
select(flights, contains("TIME"))
```

```{r}
# to change how select deals with the case default
select(flights, starts_with("TIME"))

# This line will be case sensitive - but it results in an error because there is no TIME column
# select(flights, match("TIME"))
```

#### 5.5 Add new variables with mutate()

mutate() always adds new columns at the end of your dataset 
see all the columns is View()

```{r}
# flights_sml is a subset of flights that has some of the variables from flights (from year to day // including month, all columns that end with delay, distance, and airtime)
flights_sml <- select(flights, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)

# adds 2 cols?; gain and speed
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)
```


##### 5.5.1 filter rows with filter()

##### 5.5.1 filter rows with filter()

#### 5.5 filter rows with filter()

##### 5.5.1 filter rows with filter()

##### 5.5.1 filter rows with filter()

##### 5.5.1 filter rows with filter()

##### 5.5.1 filter rows with filter()

##### 5.5.1 filter rows with filter()

##### 5.5.1 filter rows with filter()

##### 5.5.1 filter rows with filter()

#### 5.5 filter rows with filter()

##### 5.5.1 filter rows with filter()


## Text Mining with R

### Ch 2: Sentient Analysis with tidy Data

#### 2.1 The sentiments dataset

```{r}
sentiments
```


### Ch 3: Analyzing word and document frequency: tf-idf

Examples from this chapter were already done in the previous homework assignment
